---
output: rmarkdown::tufte_handout
---

Narrative: an Introduction
========================================================

*A simple introductory analysis using the Narrative package.*

```{r echo=FALSE, message=FALSE}
if(!require(ggplot2)){
  install.packages("ggplot2")
}
if(!require(tm)){
  install.packages("tm")
}

# working directory should already be set...
wd<-"C://dev//Datalab//Narrative"
setwd(wd)
```

First we wish to read some sample documents in to the corpus

```{r}
corp<-Narrative::readSeparateText(paste(wd,"//inst//extdata//Sample Docs (txt) - Header",sep=""))
```

```{r echo=FALSE}
# create term document matrix from corpus
#tdm <- TermDocumentMatrix(corp)
#dtm <- DocumentTermMatrix(corp)
#dtm<-removeSparseTerms(dtm,sparse=0.4)
```

The `readSeparateText` functions create a Corpus object from the files in the given directory, changing the name of each document to the filenames. We can then apply some cleaning function to the corpus from the *tm* package, using the `content_transformer` function to ensure the transformations are applied only to the text of the document and not the associated meta data. First we remove all non-ASCII characters:
```{r}
  clean_non_ascii <- function(x){iconv(x, "latin1", "ASCII", sub="")}
  corp<-tm_map(corp,content_transformer(clean_non_ascii))
```

Then remove punctuation:
```{r}
  corp<-tm_map(corp,content_transformer(removePunctuation))
  #corp<-tm_map(corp,content_transformer(removePunctuation),preserve_intra_word_dashes=T)
```

Remove numbers:
```{r}
  corp<-tm_map(corp,content_transformer(removeNumbers))
```

Eliminate extra whitespace (multiple whitespace characters are collapsed to a single blank):
```{r}
  corp<-tm_map(corp,content_transformer(stripWhitespace))
```

Convert all characters in the documents to lower case:
```{r}  
  corp<-tm_map(corp,content_transformer(tolower))
```

Remove stopwords, using the a default english stopword library supplied with *tm*:
```{r}  
  corp<-tm_map(corp,removeWords,stopwords("english"))
```

Stem all of the words in the document to their base:
```{r}  
  corp<-tm_map(corp,stemDocument,language="english")
```

Our corpus of documents is now cleaned and stemmed. We now wish to construct a dictionary to compare against. In this case we load both positive and negative dictionaries in to a single corpus document, and apply the samecleaning and stemming techniques as to the corpus across the whole dictionary corpus.

First, read the positive dictionary:

```{r echo=TRUE}
positive_dictionary<-Narrative::readSeparateText(paste(wd,"//inst//extdata//Dictionaries//Financial//positive",sep=""))
meta(positive_dictionary[[1]],tag="heading")<-"positive"
```

Add negative dictionary to corpus:
```{r}
negative_dictionary<-Narrative::readSeparateText(paste(wd,"//inst//extdata//Dictionaries//Financial//negative",sep=""))
meta(negative_dictionary[[1]],tag="heading")<-"negative"

dictionaries<-c(positive_dictionary,negative_dictionary)
```

Transform dictionaries, using the `transform_docs` function that combines all of the transformations described above in to a single function call:
```{r}
dictionaries<-Narrative::transform_docs(dictionaries)
```

We can then use these dictionaries to calculate the sentiment of each document in our corpus. Here we use a simple word counting approach that sums the number of positive and negative words, calculates the difference and divides by the number of words in the document:
```{r, warning=FALSE}
corp<-Narrative::corpusSentiment(corp
                                ,dict.positive=dictionaries[meta(dictionaries,"heading")=='positive']
                                ,dict.negative=dictionaries[meta(dictionaries,"heading")=='negative']
                                ,normalisation="character"
                                )
```

The plot below shows the relative sentiments of all the documents:
```{r, warning=FALSE, echo=TRUE}
qplot(seq_along(meta(corp,tag="sentiment")), sort(unlist(meta(corp,tag="sentiment"))),geom="bar",stat="identity",xlab="doc",ylab="sentiment score")
```

